---
title: "assignment2"
output: html_document
---

| B | 2.2b, 2.3a, 2.5a | 3.10 | Statistics: ANCOVA (type+diameter+height); Optimization: base diet LP; staffing (covering). |

Exercise 2.2 Trees
The ‘Amsterdamsche Bos’ Forestry wishes to estimate the total wood volume of the trees on
its domain. To this end the Forestry has cut a sample of 59 trees of their most prevalent type
‘Beech’ and ‘Oak’. The volume of these trees alongside with their height and trunk diameter
have been measured. The latter two characteristics can be measured in the field without
sacrificing the tree. The Forestry hypothesizes that these are predictive of the tree’s volume.
The data (in the file treeVolume.txt) from the cut trees is to uncover this relationship, taking
into account the tree type.
a) (0.7) Investigate whether tree type influences volume by performing ANOVA, without
taking the diameter or height into account. Is an oak more voluminous than a beech?
What are the estimated volumes for the two tree types?
```{r}
data = read.table("treeVolume.txt", header = TRUE)
data$type <- factor(data$type)
model_a = lm(volume ~ type, data = data)

summary(model_a)
anova(model_a)

```
There is no statistically significant difference in average volume between Oak and Beech trees in this sample.
That means, based on these 59 trees, we cannot conclude that Oaks are more voluminous than Beeches.

A one-way ANOVA comparing mean wood volume between Beech and Oak trees showed no significant difference (F = 1.90, p = 0.17).
The estimated mean volumes were ≈ [insert your group means]; hence, although Oaks appear slightly larger on average, this difference is not statistically significant at the 5 % level.

b) (1.1) Investigate whether the tree type influences the volume, now including diameter
and height (but no interactions) as predictors into the analysis, which results in an
ANCOVA analysis. How does tree type influence volume? What are the estimated
volumes for the two tree types with average diameter and height? (Hint: use function
predict.) How does diameter influence volume? Graphically (or propose and perform
a relevant test) investigate whether this dependence is similar for both tree types.

```{r}
data = read.table("treeVolume.txt", header=TRUE)
data$type = factor(data$type)

model_b <- lm(volume ~ diameter + height + type, data = data)

summary(model_b)
anova(model_b)          
```
2.2B NOT YET DONE!!!!!!!!!!!!!!!!!!!!!!!!

