---
title: "assignment2"
output: html_document
---

| B | 2.2b, 2.3a, 2.5a | 3.10 | Statistics: ANCOVA (type+diameter+height); Optimization: base diet LP; staffing (covering). |

| C | 2.2a, 2.2c, 2.4b, 2.5b, 2.1d | 3.40 | Statistics: ANOVA + transformation; Optimization: fixed-charge transport; L1 staffing fit; prediction intervals. |

Exercise 2.1 Scholastic Aptitude Test
The data in the file sat.txt concerns data on the Scholastic Aptitude Test (SAT) for pupils in
the United States in 1994/1995. The column expend contains the mean expenses per pupil (in
$ per pupil), ratio is the pupil/teacher ratio, salary is the mean salary of teachers, takers
is the percentage of pupils that takes the SAT. Variables verbal and math are partial scores
of the total SAT score in total and not used in the analysis. All data are per state.


Exercise 2.2 Trees
The ‘Amsterdamsche Bos’ Forestry wishes to estimate the total wood volume of the trees on
its domain. To this end the Forestry has cut a sample of 59 trees of their most prevalent type
‘Beech’ and ‘Oak’. The volume of these trees alongside with their height and trunk diameter
have been measured. The latter two characteristics can be measured in the field without
sacrificing the tree. The Forestry hypothesizes that these are predictive of the tree’s volume.
The data (in the file treeVolume.txt) from the cut trees is to uncover this relationship, taking
into account the tree type.
a) (0.7) Investigate whether tree type influences volume by performing ANOVA, without
taking the diameter or height into account. Is an oak more voluminous than a beech?
What are the estimated volumes for the two tree types?
```{r}
data = read.table("treeVolume.txt", header = TRUE)
data$type <- factor(data$type)
model_a = lm(volume ~ type, data = data)

summary(model_a)
anova(model_a)

```
There is no statistically significant difference in average volume between Oak and Beech trees in this sample.
That means, based on these 59 trees, we cannot conclude that Oaks are more voluminous than Beeches.

A one-way ANOVA comparing mean wood volume between Beech and Oak trees showed no significant difference (F = 1.90, p = 0.17).
The estimated mean volumes were ≈ [insert your group means]; hence, although Oaks appear slightly larger on average, this difference is not statistically significant at the 5 % level.

b) (1.1) Investigate whether the tree type influences the volume, now including diameter
and height (but no interactions) as predictors into the analysis, which results in an
ANCOVA analysis. How does tree type influence volume? What are the estimated
volumes for the two tree types with average diameter and height? (Hint: use function
predict.) How does diameter influence volume? Graphically (or propose and perform
a relevant test) investigate whether this dependence is similar for both tree types.

```{r}
data = read.table("treeVolume.txt", header=TRUE)
data$type = factor(data$type)

model_b <- lm(volume ~ diameter + height + type, data = data)

summary(model_b)
anova(model_b)          
```




Exercise 2.3 Optimal Product Mix
Consider the prices and nutrients of the following foods.
You are interested in finding the cheapest diet that satisfies the following (you can take
fractions of servings):
Calories must be at least 2000,
Fat must be at least 50g,
Protein must be at least 100g,
Carbohydrates must be at least 250g.


Exercise 2.4 Transportation Problem (will be discussed in Lecture 9)
Consider a transportation problem in which goods have to be transported from 3 sources
labeled as S1, S2 and S3 to 4 destinations labeled as D1, D2, D3 and D4. The costs in euros
per unit transported from a source to a destination are given in the following table. It also
contains the demand at the destinations and the available supply at the sources:


Exercise 2.5 Call Center Staffing
Based on predictive analytics and some service level requirement, the required staffing in a
call center from 09:00 to 21:00 is given Table 1. There are 2 types of shifts available:
(i) 8-hour working shifts with 30 minutes unpaid break after 4 working hours (note: this
shift takes 8.5 hours in total). These shifts can start every half an hour from 09:00 to
12:30. The wage is 20 euro/hour.
(ii) 4-hour working shifts without a break that can start every half an hour from 09:00 to
17:00. The wage is 24 euro/hour.